Containers and R---the Rockerverse and beyond

### 2020-01-10 manuscript.Rmd
move section about non-Docker tools up

# Container Images

>>>>

Docker is not the only containerisation software.
An alternative stemming from the domain of high-perfomance computing is **Singularity** \citep{kurtzer_singularity_2017}.
Singularity can run Docker images, and in the case of Rocker works out of the box if the main process is R, e.g., in `rocker/r-base`, but does not succeed in running images where there is an init script, e.g., in containers that by default run RStudio.
In the latter case, a `Singularity` file, a recipe akin to a `Dockerfile`, needs to be used.
To date, no comparable image stack to Rocker exists on [Singularity Hub](https://singularity-hub.org/).
A further tool for running containers is [**podman**](https://github.com/containers/libpod), which also can build `Dockerfile`s and run Docker images.
Proof of concepts for using podman to build and run Rocker containers exist\footnote{See \href{https://github.com/nuest/rodman}{https://github.com/nuest/rodman} and \href{https://github.com/rocker-org/rocker-versioned/issues/187}{https://github.com/rocker-org/rocker-versioned/issues/187}}.
Yet the prevelance of Docker, especially in the broader user community beyond experts or niche systems, and the vast amount of blog posts and courses for Docker, currently caps specific development efforts for both Singularity and podman in the R community.
This might quickly change when usability and spread increase, or security features such as rootless/unprivileged containers, which both these tools support out of the box, become more sought after.

# Container images
